---
title: 'Ohio Covid-19 : Can We Trust the Numbers?'
author: "IllicitDavid Jackson"
date: "2/12/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,message = FALSE,error=FALSE)
```

# Is The Trend Still Our Firend?
* Can we Trust Ohio's Covid-19 Cases and Deaths
* Ohio health dept. to restructure after 4K unreported deaths(AP)
* https://wpta21.com/2021/02/12/governor-expected-to-discuss-underreported-covid-19-deaths/

```{r}
library(ggplot2)
library(scales)
library(RSQLite)
library(dplyr)
```
```{r}
rm(list=ls())
db <- db <- dbConnect(RSQLite::SQLite(),dbname= "../COVID-19-DB/NYTimes.sqlite3")
df <- dbGetQuery(db,"select * from STATESDAILY")
df <- df[order(df$date),]
df$date <- as.Date(df$date)
oh <- subset(df,state =="Ohio" & date >="2020-03-10" )

summary(oh$new_cases,na.rm=TRUE)
summary(oh$new_deaths,na.rm=TRUE)
library(benford.analysis)
library(pracma)  ## Inclused a funcion to do various rolling means calc.
```

## Comparison of Ohio Cases and Deaths

```{r}
oh$mad7 <- movavg(oh$new_deaths,7, type="e")
oh$mad28 <- movavg(oh$new_deaths,28, type="e")

ggplot(oh) + geom_line(aes(x=date,y=mad7)) +
  labs(title="Ohio Deaths: 7 Day Moving Average") +
  ylim(0,90)

ggplot(oh) + geom_line(aes(x=date,y=mad28)) +
  labs(title="Ohio Deaths: 28 Day Moving Average") +
  ylim(0,90)

```

# Benford's Analysis: benford.analysis packager
## Ohio Daily Deaths: First and First Two Digits

```{r}
oh_cases <- oh %>% select(new_cases)
oh_deaths <- oh %>% select(new_deaths)

bl_deaths <- benford(oh_deaths$new_deaths,1)
plot(bl_deaths,multiple = FALSE,select="digits")
#
bl_deaths2 <- benford(oh_deaths$new_deaths,2)
plot(bl_deaths2,multiple = FALSE,select="digits")
```

## Comparing New York State Covid-19 Deaths
* Source: https://www.nytimes.com/2021/01/28/nyregion/nursing-home-deaths-cuomo.html
* Quote,"N.Y. Severely Undercounted Virus Deaths in Nursing Homes,(AG)


```{r}
new_york <- subset(df,state =="New York")
ny_cases <- new_york %>% select(new_cases)
ny_deaths <- new_york %>% select(new_deaths)

```

```{r}
new_york$mad7 <- movavg(new_york$new_deaths,7, type="e")
new_york$mad28 <- movavg(new_york$new_deaths,28, type="e")

ggplot(new_york) + geom_line(aes(x=date,y=mad7)) +
  labs(title="New York Deaths: 7 Day Moving Average") 
  

ggplot(new_york) + geom_line(aes(x=date,y=mad28)) +
  labs(title="New York Deaths: 28 Day Moving Average")

```
## New York State's Daily Deaths: First and First Two Digits

```{r}
bl_deaths <- benford(ny_deaths$new_deaths,1)
plot(bl_deaths,multiple = FALSE,select="digits")
#
bl_deaths2 <- benford(ny_deaths$new_deaths,2)
plot(bl_deaths2,multiple = FALSE,select="digits")
```


